---
title: "Proteomics analysis TQ vs. NPA"
author: "Fernando Garrido Muñoz"
date: "`r format(Sys.time(), '%a %d, %b, %Y')`"
output: html_document
---

# Load required packages

```{r, warning=FALSE, results='hide', message=FALSE}

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
library(tidyverse)
library(stringr)
library(readxl)
library(writexl)
library(ggplot2)
library(gplots)
library(VennDetail)
library(kableExtra)
library(gtsummary)
library(circlize)
library(ComplexHeatmap)
library(gprofiler2)

```

# Load and pre-process proteomic data

```{r}

data <- read_excel("TAvsNPA_OPT.xlsx")
data <- as.data.frame(data)

```

## Filter rows which contain NAs

```{r}

# Remove duplicate rows based on Protein.IDs:(no duplicated rows)

data <- data[!duplicated(data$Genes), ]

# Remove rows with NAs protein IDs

data <- data[!is.na(data$Protein.Ids),]

```

## Rework data frame to select our samples with all protein ids (rows):

```{r}

sample_columns = seq(4,length(data),1)
df_protein <- data[, sample_columns]
rownames(df_protein) <- data$Genes
df_protein[,c(1:10)] <- apply(df_protein[,c(1:10)],2,
                        function(x) as.numeric(as.character(x)))

# We must extract only the samples we need 

# Which is the % of NAs?

sum(is.na(df_protein)/prod(dim(df_protein))) *100

# 18.52 %

```

## Order columns in groups:

### ALL samples

```{r}

samples <- read_excel("inventario_post_procesamiento_TQ.xlsx", 
                      sheet = "SamplesConditions")
samples <- as.data.frame(samples)
samples$jensen_condition <- as.factor(samples$jensen_condition) # turn this data to factor
# class(samples$jensen_condition) this is used to check if it has correctly changed to factor
samples$type <- as.factor(samples$type)
# order_samples <- samples$jensen_condition
rownames(samples) <- samples$new_sample

# Add a more specific column with condition + type:

samples <- samples %>% mutate(cond_type=case_when(jensen_condition=="control" & type=="NF" ~ "NPA Control",
                                       jensen_condition=="control" & type=="TQ" ~ "TA Control",
                                       jensen_condition=="bpd" & type=="NF" ~ "NPA BPD",
                                       jensen_condition=="bpd" & type=="TQ" ~ "TA BPD",
                                       jensen_condition=="fallecido" & type=="NF" ~ "NPA Deceased",
                                       jensen_condition=="fallecido" & type=="TQ" ~ "TA Deceased"),
                              cod_centro=case_when(hospital_abrev=="Cádiz" ~ "01",
                                                   hospital_abrev=="Basurto" ~ "02",
                                                   hospital_abrev=="Vigo" ~ "03",
                                                   hospital_abrev=="Girona" ~ "04",
                                                   hospital_abrev=="Clinic" ~ "05",
                                                   hospital_abrev=="HGM" ~ "06",
                                                   hospital_abrev=="León" ~ "07"))

samples$cond_type <- as.factor(samples$cond_type)
samples <- samples[samples$new_sample %in% colnames(df_protein),]
samples <- samples[order(samples$patient),]
samples$new_ID <- c("NPA_S4","TA_S4","NPA_S1", "TA_S1",
                    "NPA_S2","TA_S2","NPA_S3", "TA_S3",
                    "NPA_S5","TA_S5") 
#samples <- samples %>%
#  mutate(def.id=paste0("Sample_",samples$new_ID))
samples$type <- factor(samples$type, levels = c("NF", "TQ"), labels = c("NPA","TA"))
# before order, we must keep only the samples we'll need in this experiment:
## NEW ADDED FEATURE TO THE HOLE SCRIPT, THIS IS VERY IMPORTANT, WE DONT NEED FURTHER
## EDITION OF THE PACIENTES XLS RESULTS TABLE IN EXCEL

df_new_prot <- df_protein[,colnames(df_protein) %in% samples$new_sample]
samples <- samples[str_order(sapply(str_extract_all(samples$new_ID, "\\d+"), 
                            tail, 1), numeric = TRUE),]
order <- samples$new_samples
df_new_prot <- df_new_prot[,order]
colnames(df_new_prot) <- samples$new_ID
samples$sample_num <- c("S1","S1","S2","S2","S3","S3",
                        "S4","S4","S5","S5")
protein.stats <- samples[,c(10:11,7)]
rownames(protein.stats) <- protein.stats$new_ID
num.proteins= c()
for (i in protein.stats$new_ID){
  h= length(df_new_prot[,i][!is.na(df_new_prot[,i])])
  num.proteins= append(num.proteins,h)
}
protein.stats$num.proteins <- num.proteins

```

# Clinical variables table:

```{r}

variables <- read_xlsx("./pacientes.xlsx")
variables <- variables[,c(3,5:6,8:9,11,54,70,72)]
variables <- merge(samples, variables, by=c("patient", "cod_centro"))
variables <- variables[!duplicated(variables$patient),]
variables$`Fecha de Nacimiento` <- as.Date(variables$`Fecha de Nacimiento`, "%d/%m/%Y")
variables$`Fecha de alta` <- as.Date(variables$`Fecha de alta`, "%d/%m/%Y")
variables$Éxitus <- factor(variables$Éxitus, levels= c("SÍ", "NO"), 
                           labels = c("Yes", "No")) 
variables <- variables %>%
  mutate(Condition=case_when(jensen_condition=="bpd" ~ "BPD",
                             jensen_condition=="control" ~ "Control",
                             jensen_condition=="fallecido" ~ "Deceased"),
         BPD.Grade=case_when(jensen=="grado 1" ~ "Grade 1",
                             jensen=="grado 2" ~ "Grade 2",
                             jensen=="grado 3" ~ "Grade 3",
                             jensen=="control" ~ "Control",
                             jensen=="fallecido" ~ "Deceased"),
         Days.hospitalization=difftime(`Fecha de alta`,`Fecha de Nacimiento`, 
                                       units = "days"),
         New.Sex=case_when(Sexo=="Mujer" ~ "Yes",
                           Sexo=="Varón" ~ "No"))  %>%
  mutate_all(function(x) gsub("Varón","Male",x)) %>%
  mutate_all(function(x) gsub("Mujer","Female",x))

variables <- variables %>%
  mutate_at(c("EG al nacimiento (semanas)", "Peso al nacimiento", 
              "Days.hospitalization", "Días 1 sem de VM"), as.numeric)
variables$Condition <- factor(variables$Condition, levels=c("Control", "BPD", "Deceased"))
variables$BPD.Grade <- factor(variables$BPD.Grade, levels=c("Control", "Grade 1",
                                                            "Grade 2", "Grade 3",
                                                            "Deceased"))
variables$New.Sex <- factor(variables$New.Sex, levels = c("Yes", "No"))
library(openxlsx)
wide.modality <- read.xlsx("wide_060324_mod.xlsx")
wide.modality <- wide.modality %>%
  mutate(Modality=case_when(Modalidad=="AC/VG" | Modalidad=="PC" | 
                              Modalidad=="PSV" | Modalidad=="PSV+VG" |
                              Modalidad=="SIMV+VG" | Modalidad=="VAFO" |
                              Modalidad=="VCRP" | Modalidad=="VM A/C" |
                              Modalidad=="VM A/C CON VG" | 
                              Modalidad=="VMC" | Modalidad=="VMC A/C" ~ "IMV",
                            Modalidad=="VNI" | Modalidad=="DUOPAP" |
                              Modalidad=="IPPVn" ~ "NIV",
                            Modalidad=="CPAP" | Modalidad=="CPAPn" ~ "nCPAP",
                            Modalidad=="Alto flujo" | Modalidad=="GNAF" |
                              Modalidad=="OAF" | Modalidad=="HFNC" ~ "HFNC",
                            Modalidad=="GNBF" ~ "LFNC",
                            Modalidad=="No O2" | 
                              Modalidad=="Sin soporte" ~ "No respiratory support"
                            ))
wide.modality$Modality <- factor(wide.modality$Modality, levels = c("IMV"))
new.variables <- merge(variables, wide.modality, by=c("patient","cod_centro"))

new.variables <- new.variables[,c(22,14:16,18,19,24,20:21)]

# View table:

new.variables %>%
  tbl_summary(
    by= Condition,
    statistic = list(all_continuous() ~ "{mean} ({sd})",        
                     BPD.Grade ~ "{n} ({p}%)",
                     New.Sex ~ "{n} ({p}%)"),   
    digits = all_continuous() ~ 1,                            
    type   = list(all_categorical() ~ "categorical",
                  `EG al nacimiento (semanas)` ~ "continuous",
                  `Días 1 sem de VM` ~ "continuous",
                  Modality ~ "categorical",
                  `Peso al nacimiento` ~ "continuous",
                  Days.hospitalization ~ "continuous"),
    label = list(New.Sex ~ "Sex, Female",
                 `EG al nacimiento (semanas)` ~ "GA (weeks)",
                 `Peso al nacimiento` ~ "Weight",
                 `Días 1 sem de VM` ~ "Days of VM at 1 week",
                 BPD.Grade ~ "BPD grade",
                 Days.hospitalization ~ "Days of hospitalization",
                 Modality ~ "Respiratory modality al 1 week")) %>%
  modify_header(label ~ "**Condition**") %>%
  modify_caption("**Table 1. Patient Clinical Characteristics**") %>%
  add_overall() %>%
  italicize_levels() %>%
  as_gt() %>%
  gt::tab_source_note(gt::md("*GA: Gestational Age; BPD: Bronchopulmonary Dysplasia;IVM: Invasive Mechanical Ventilation*"))

# Save table as .docx

new.variables %>%
  tbl_summary(
    by= Condition,
    statistic = list(all_continuous() ~ "{mean} ({sd})",        
                     BPD.Grade ~ "{n} ({p}%)",
                     New.Sex ~ "{n} ({p}%)"),   
    digits = all_continuous() ~ 1,                            
    type   = list(all_categorical() ~ "categorical",
                  `EG al nacimiento (semanas)` ~ "continuous",
                  `Días 1 sem de VM` ~ "continuous",
                  Modality ~ "categorical",
                  `Peso al nacimiento` ~ "continuous",
                  Days.hospitalization ~ "continuous"),
    label = list(New.Sex ~ "Sex, Female",
                 `EG al nacimiento (semanas)` ~ "GA (weeks)",
                 `Peso al nacimiento` ~ "Weight",
                 `Días 1 sem de VM` ~ "Days of VM at 1 week",
                 BPD.Grade ~ "BPD grade",
                 Days.hospitalization ~ "Days of hospitalization",
                 Modality ~ "Respiratory modality al 1 week")) %>%
  modify_header(label ~ "**Condition**") %>%
  modify_caption("**Table 1. Patient Clinical Characteristics**") %>%
  add_overall() %>%
  italicize_levels() %>%
  as_gt() %>%
  gt::tab_source_note(gt::md("*GA: Gestational Age; BPD: Bronchopulmonary Dysplasia;IVM: Invasive Mechanical Ventilation*")) %>%
  gt::gtsave(filename = "Table_1_modality.docx")

```


# Number of proteins per sample

```{r}

bar.prots <- ggplot(protein.stats, aes(x = sample_num, y = num.proteins, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(labels = c("NPA", "TA"),
                    values=c("#1A85FF", "#D41159"),
                    name="") +
  geom_text(aes(label = num.proteins), vjust = -0.8,
            size=3, fontface="bold.italic",
            position = position_dodge(.9),
            colour = "black") +
  ylim(0, 3000) +
  ylab("Number of identified proteins") +
  xlab("Samples")

ggsave(plot = bar.prots, file="bar_prots.svg", width = 9)
ggsave(plot = bar.prots, file="bar_prots.pdf", width = 9)

```

# Mean difference in protein number betweeen samples

```{r}

difference.prot <- c()
for (i in 1:10){
  dif= abs(protein.stats$num.proteins[i]-protein.stats$num.proteins[i+1])
  difference.prot= append(difference.prot,dif)
}
difference.prot <- difference.prot[c(1,3,5,7,9)]
mean(difference.prot) # 610
median(difference.prot)
quantile(difference.prot, probs = c(0,0.25,0.50,0.75,1))
sd(difference.prot) # +- 386

```

# Log transformation of data

```{r}

protein.log <- log2(df_new_prot)

```


# Comparative Analysis: NPA vs. TA (common and specific proteins)

```{r}
library(svglite)
TA.prot <- protein.log[,samples$new_ID[samples$type=="TA"]]
sum(is.na(TA.prot)/prod(dim(TA.prot))) *100 # 44.39 NA
TA.prot <- TA.prot[!rowSums(is.na(TA.prot)),] # 1148 prots
NPA.prot <- protein.log[,samples$new_ID[samples$type=="NPA"]]
sum(is.na(NPA.prot)/prod(dim(NPA.prot))) *100 # 43.98 NA
NPA.prot <- NPA.prot[!rowSums(is.na(NPA.prot)),] # 826 prots

comparative <- venndetail(list(NPA=rownames(NPA.prot), TA=rownames(TA.prot)))
comp.res <- result(comparative)
svglite::svglite("Figures/venn.svg")
plot(comparative)
dev.off()

exclusive.TA <- TA.prot[comp.res$Detail[comp.res$Subset=="TA"],]
exclusive.NPA <- NPA.prot[comp.res$Detail[comp.res$Subset=="NPA"],]

```

# Heatmap of top 100 proteins NPA and TA

```{r}

means.excl.TA <- rowMeans(exclusive.TA)
top100.TA <- means.excl.TA[order(means.excl.TA, decreasing=T)]
top100.TA <- names(top100.TA[1:100])
means.excl.NPA <- rowMeans(exclusive.NPA)
top100.NPA <- means.excl.NPA[order(means.excl.NPA, decreasing=T)]
top100.NPA <- names(top100.NPA[1:100])

```

# Heatmap for TA

```{r}

# Extract final top 50 proteins matrix

mat.top100.TA <- as.matrix(exclusive.TA[top100.TA[1:50],])
mat.top100.TA.sca <- as.matrix(t(apply(mat.top100.TA, 1,scale)))
colnames(mat.top100.TA.sca) <- colnames(mat.top100.TA)
means.TA <- as.matrix(means.excl.TA[top100.TA[1:50]])
colnames(means.TA) <- "AveExp"
col_AveExpr <- colorRamp2(c(quantile(means.TA)[1],
                            quantile(means.TA)[4]),
                          c("white", "red"))
col.fun <- colorRamp2(c(-2, 2), c("white", "red"))

ha <- HeatmapAnnotation(summary = anno_summary(gp=gpar(fill="red"), 
                                               height =unit(1, "cm")))

h1 <- Heatmap(mat.top100.TA.sca, cluster_rows = F, 
              column_labels = colnames(mat.top100.TA), 
              name="Z-score", cluster_columns = F,
              width = unit(7,"cm"),
              show_row_names = F,
              col=col.fun)
h2 <- Heatmap(means.TA, row_labels = rownames(mat.top100.TA), 
              cluster_rows = F, 
              name= "AveExpr", 
              col = col_AveExpr, 
              cell_fun = function(j,i,x,y,w,h,col){
                grid.text(round(means.TA[i,j],2),x,y,
                          gp=gpar(fontsize=4, fontface="bold"))},
              width = unit(1, "cm"),
              top_annotation = ha,
              row_names_gp= gpar(fontsize=4),
              row_names_side = "left")

h <- h2 + h1
h
svglite::svglite("Figures/heatmap_TA.svg")
h
dev.off()


```

# Heatmap for NPA

```{r}

# Extract final top 50 proteins matrix

mat.top100.NPA <- as.matrix(exclusive.NPA[top100.NPA[1:50],])
mat.top100.NPA.sca <- as.matrix(t(apply(mat.top100.NPA, 1,scale)))
colnames(mat.top100.NPA.sca) <- colnames(mat.top100.NPA)
means.NPA <- as.matrix(means.excl.NPA[top100.NPA[1:50]])
colnames(means.NPA) <- "AveExp"
col_AveExpr_NPA <- colorRamp2(c(quantile(means.NPA)[1],
                            quantile(means.NPA)[4]),
                          c("white", "blue"))
col.fun.NPA <- colorRamp2(c(-2, 2), c("white", "blue"))

ha.NPA <- HeatmapAnnotation(summary = anno_summary(gp=gpar(fill="blue"), 
                                               height =unit(1, "cm")))

h1.NPA <- Heatmap(mat.top100.NPA.sca, cluster_rows = F, 
              column_labels = colnames(mat.top100.NPA), 
              name="Z-score", cluster_columns = F,
              width = unit(7,"cm"),
              show_row_names = F,
              col=col.fun.NPA)
h2.NPA <- Heatmap(means.NPA, row_labels = rownames(mat.top100.NPA), 
              cluster_rows = F, 
              name= "AveExpr", 
              col = col_AveExpr_NPA, 
              cell_fun = function(j,i,x,y,w,h,col){
                grid.text(round(means.NPA[i,j],2),x,y,
                          gp=gpar(fontsize=4, fontface="bold"))},
              width = unit(1, "cm"),
              top_annotation = ha.NPA,
              row_names_gp= gpar(fontsize=4),
              row_names_side = "right")

h.NPA <- h1.NPA + h2.NPA 
h.NPA

svglite::svglite("Figures/heatmap_NPA.svg")
h.NPA
dev.off()

```

# Heatmap with all proteins with all samples

```{r}

noNA.prot <- protein.log[!rowSums(is.na(protein.log)),]
mat.all <- as.matrix(t(apply(noNA.prot, 1,scale)))
colnames(mat.all) <- colnames(noNA.prot)
col.fun.all <- colorRamp2(c(-4, 0, 4), c("blue","white", "red"))
h.all <- Heatmap(mat.all, cluster_rows = T, 
              column_labels = colnames(mat.all), 
              name="Z-score", cluster_columns = T,
              width = unit(8,"cm"),
              show_row_names = F,
              col=col.fun.all,
              show_row_dend = F,
              clustering_method_columns = "ward.D")
h.all
svglite::svglite("Figures/heatmap_ALL.svg")
h.all
dev.off()

```

# Scatterplots for each sample:

```{r}
# Sample 1

plot(noNA.prot[,1],noNA.prot[,2],
     pch=19,col="grey",xlab="NPA_1",ylab="TA_1",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 1")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,1],noNA.prot[,2]),
                  digits = 2)), collapse=""))

svglite::svglite("Figures/scatter_S1.svg")
plot(noNA.prot[,1],noNA.prot[,2],
     pch=19,col="grey",xlab="NPA_1",ylab="TA_1",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 1")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,1],noNA.prot[,2]),
                  digits = 2)), collapse=""))
dev.off()

cor.test(noNA.prot[,1],noNA.prot[,2])

# Sample 2

plot(noNA.prot[,3],noNA.prot[,4],
     pch=19,col="grey",xlab="NPA_2",ylab="TA_2",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 2")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,3],noNA.prot[,4]),
                  digits = 2)), collapse=""))

svglite::svglite("Figures/scatter_S2.svg")
plot(noNA.prot[,3],noNA.prot[,4],
     pch=19,col="grey",xlab="NPA_2",ylab="TA_2",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 2")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,3],noNA.prot[,4]),
                  digits = 2)), collapse=""))
dev.off()

cor.test(noNA.prot[,3],noNA.prot[,4])

# Sample 3

plot(noNA.prot[,5],noNA.prot[,6],
     pch=19,col="grey",xlab="NPA_3",ylab="TA_3",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 3")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,5],noNA.prot[,6]),
                  digits = 2)), collapse=""))

svglite::svglite("Figures/scatter_S3.svg")
plot(noNA.prot[,5],noNA.prot[,6],
     pch=19,col="grey",xlab="NPA_3",ylab="TA_3",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 3")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,5],noNA.prot[,6]),
                  digits = 2)), collapse=""))
dev.off()

cor.test(noNA.prot[,5],noNA.prot[,6])

# Sample 4

plot(noNA.prot[,7],noNA.prot[,8],
     pch=19,col="grey",xlab="NPA_4",ylab="TA_4",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 4")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,7],noNA.prot[,8]),
                  digits = 2)), collapse=""))

svglite::svglite("Figures/scatter_S4.svg")
plot(noNA.prot[,7],noNA.prot[,8],
     pch=19,col="grey",xlab="NPA_3",ylab="TA_3",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 3")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,7],noNA.prot[,8]),
                  digits = 2)), collapse=""))
dev.off()

cor.test(noNA.prot[,7],noNA.prot[,8])

# Sample 5

plot(noNA.prot[,9],noNA.prot[,10],
     pch=19,col="grey",xlab="NPA_5",ylab="TA_5",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 5")
text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,9],noNA.prot[,10]),
                  digits = 2)), collapse=""))

svglite::svglite("Figures/scatter_S5.svg")
plot(noNA.prot[,9],noNA.prot[,10],
     pch=19,col="grey",xlab="NPA_3",ylab="TA_3",cex=0.5,
     xlim=c(10,20), ylim=c(10,20), main="Sample 3")

text(x=18,y=12,
     labels = paste(c(
              "R2 = ",
              round(cor(noNA.prot[,9],noNA.prot[,10]),
                  digits = 2)), collapse=""))
dev.off()

cor.test(noNA.prot[,9],noNA.prot[,10])

```

# Enrichment analysis for Exclusive NPA and TA proteins

```{r}

enrich.NPA <- rownames(exclusive.NPA)
enrich.TA <- rownames(exclusive.TA)

```

## NPA and TA 

```{r}

res.enr.NPA <- gost(query = enrich.NPA, 
                organism = "hsapiens", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "fdr", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = NULL, as_short_link = FALSE)

gostplot(res.enr.NPA, capped = TRUE, interactive = TRUE)

result.table.NPA <- res.enr.NPA$result
publish_gosttable(res.enr.NPA, highlight_terms = res.enr.NPA$result[c(1:4),],
                        use_colors = TRUE, 
                        show_columns = c("source", "term_name", "term_size",
                                         "intersection_size"),
                        filename = NULL)

res.all <- gost(query = list("NPA" = enrich.NPA,
                             "TA" = enrich.TA, 
                       multi_query = F))#,
                #sources = c("GO:BP", "GO:CC", "GO:MF"))
svglite::svglite("Figures/enrich_all_groups.svg")
gostplot(res.all, capped = TRUE, interactive = F)
dev.off()
result.table.all <- res.all$result
r.tabl.all.NPA <- result.table.all[result.table.all$query=="NPA",]
r.tabl.all.TA <- result.table.all[result.table.all$query=="TA",]
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="TF",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="MIRNA",]))

hola <- venndetail(list(NPA=r.tabl.all.NPA$term_id, TA=r.tabl.all.TA$term_id))
hola.res <- result(hola)
shared.ids <- hola.res$Detail[hola.res$Subset=="Shared"]

red.all.NPA <- r.tabl.all.NPA[r.tabl.all.NPA$term_id %in% shared.ids,]
red.all.NPA <- red.all.NPA[,c(6,10:11)]
red.all.NPA$source <- factor(red.all.NPA$source, levels = c("GO:MF","GO:CC",
                                                            "GO:BP","REAC",
                                                            "HPA"))
red.all.NPA$intersection_size <- as.numeric(red.all.NPA$intersection_size)
red.all.TA <- r.tabl.all.TA[r.tabl.all.TA$term_id %in% shared.ids,]
red.all.TA <- red.all.TA[,c(6,10:11)]
red.all.TA$source <- factor(red.all.TA$source, levels = c("GO:MF","GO:CC",
                                                            "GO:BP","REAC",
                                                            "HPA"))
red.all.TA$intersection_size <- as.numeric(red.all.TA$intersection_size)

```

## How many terms are in each enrichment

### NPA

```{r}

length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="GO:BP",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="GO:CC",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="GO:MF",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="KEGG",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="REAC",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="HPA",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="HP",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="WP",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="CORUM",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="TF",]))
length(rownames(r.tabl.all.NPA[r.tabl.all.NPA$source=="MIRNA",]))

```

### TA

```{r}

length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="GO:BP",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="GO:CC",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="GO:MF",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="KEGG",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="REAC",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="HPA",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="HP",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="WP",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="CORUM",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="TF",]))
length(rownames(r.tabl.all.TA[r.tabl.all.TA$source=="MIRNA",]))
```

## Histograms for shared Enrichment functions

### NPA

```{r}

red.all.NPA <- red.all.NPA[order(desc(red.all.NPA$source),
                                 red.all.NPA$intersection_size),]
red.all.NPA$term_name <- str_to_title(red.all.NPA$term_name)
red.all.NPA$term_name <- factor(red.all.NPA$term_name, 
                                levels = red.all.NPA$term_name)
write_xlsx(x= red.all.NPA[,c(3,2,1)], "Tables/Table_2_1.xlsx")

bar.enr.NPA.450 <- ggplot(red.all.NPA, aes(x = intersection_size, 
                                       y = term_name, 
                                     fill = source)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(labels = c("GO: Mollecular Function", 
                                "GO: Cellular Component",
                                "GO: Biogical Process",
                                "REACTOME Pathway",
                               "Human Protein Atlas"),
                    values=c("#5CA896", "#2800B5",
                             "#CBA042","#71EEE2",
                             "#761769"),
                    name="Source Database") +
  geom_text(aes(label = intersection_size), hjust = -0.5,
            size=2, fontface="bold.italic",
            colour = "black") +
  xlim(0, 450) +
  xlab("Number of proteins for each term") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 5))

bar.enr.NPA.450
ggsave(plot = bar.enr.NPA.450, "bar_enr_NPA_450.svg", width = 9)
ggsave(plot = bar.enr.NPA.450, "bar_enr_NPA_450.pdf", width = 9)
ggsave(plot = bar.enr.NPA.450, "bar_enr_NPA_450.png", width = 9)


bar.enr.NPA.200 <- ggplot(red.all.NPA, aes(x = intersection_size, 
                                       y = term_name, 
                                     fill = source)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(labels = c("GO: Mollecular Function", 
                                "GO: Cellular Component",
                                "GO: Biogical Process",
                                "REACTOME Pathway",
                               "Human Protein Atlas"),
                    values=c("#5CA896", "#2800B5",
                             "#CBA042","#71EEE2",
                             "#761769"),
                    name="Source Database") +
  geom_text(aes(label = intersection_size), hjust = -0.5,
            size=2, fontface="bold.italic",
            colour = "black") +
  xlim(0, 200) +
  xlab("Number of proteins for each term") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 5))

bar.enr.NPA.200
ggsave(plot = bar.enr.NPA.200, "bar_enr_NPA_200.svg", width = 9)
ggsave(plot = bar.enr.NPA.200, "bar_enr_NPA_200.pdf", width = 9)
ggsave(plot = bar.enr.NPA.200, "bar_enr_NPA_200.png", width = 9)

```


### TA

```{r}

red.all.TA <- red.all.TA[order(desc(red.all.TA$source),
                                 red.all.TA$intersection_size),]
red.all.TA$term_name <- str_to_title(red.all.TA$term_name)
red.all.TA$term_name <- factor(red.all.TA$term_name, 
                                levels = red.all.TA$term_name)
write_xlsx(x= red.all.TA[,c(3,2,1)], "Tables/Table_2_2.xlsx")
bar.enr.TA <- ggplot(red.all.TA, aes(x = intersection_size, 
                                       y = term_name, 
                                     fill = source)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(labels = c("GO: Mollecular Function", 
                                "GO: Cellular Component",
                                "GO: Biogical Process",
                                "REACTOME Pathway",
                               "Human Protein Atlas"),
                    values=c("#5CA896", "#2800B5",
                             "#CBA042","#71EEE2",
                             "#761769"),
                    name="Source Database") +
  geom_text(aes(label = intersection_size), hjust = -0.5,
            size=2, fontface="bold.italic",
            colour = "black") +
  xlim(0, 450) +
  xlab("Number of proteins for each term") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 5))

bar.enr.TA
ggsave(plot = bar.enr.TA, "bar_enr_TA.svg", width = 9)
ggsave(plot = bar.enr.TA, "bar_enr_TA.pdf", width = 9)
ggsave(plot = bar.enr.TA, "bar_enr_TA.png", width = 9)
```



## Histograms for individual TOP enriched functions

### NPA 

```{r}

ind.all.NPA <- r.tabl.all.NPA[!r.tabl.all.NPA$term_id %in% shared.ids,]
ind.all.NPA <- ind.all.NPA[,c(3,6:7,10:11)]
ind.all.NPA$source <- factor(ind.all.NPA$source, levels = c("GO:BP","GO:CC",
                                                            "GO:MF","HPA",
                                                            "HP","REAC",
                                                            "KEGG","WP"))
ind.all.NPA.HPA <- ind.all.NPA[ind.all.NPA$source=="HPA",]
ind.all.NPA <- ind.all.NPA[!ind.all.NPA$source=="HPA",]
ind.all.NPA <- ind.all.NPA[order(ind.all.NPA$precision),]
ind.all.NPA$term_name <- str_to_title(ind.all.NPA$term_name)
ind.all.NPA$term_name[18] <- "Keratinization (II)"
ind.all.NPA$term_name <- factor(ind.all.NPA$term_name, 
                                levels = ind.all.NPA$term_name)

bar.ind.NPA <- ggplot(ind.all.NPA, aes(precision,term_name)) +
  geom_point(aes(fill= p_value, size= intersection_size,
                 color=source), shape=21, stroke=1.2) + 
  scale_color_manual(labels = c("GO: Biogical Process", 
                               "GO: Cellular Component",
                               "GO: Mollecular Function",
                               "Human Phenotype Ontology",
                               "REACTOME Pathway",
                               "KEGG database",
                               "WikiPathways"
                               ),
                    values=c("#648FFF", "#785EF0",
                             "#DC267F",
                             "#FFB000","#A50D3E",
                             "#484661","#9FCCE9"),
                    name="Source Database") +
  scale_fill_gradient(low="wheat", high= "olivedrab4",
                              name="FDR") +
  scale_size_continuous(breaks = c(10,20,30),
                        limits = c(0,35),
                        name = "Observed Gene Count") +
  xlim(0,0.20) +
  xlab("Gene Ratio") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 7))

bar.ind.NPA
ggsave(plot = bar.ind.NPA, "Enrich_plots/svg/bar_ind_NPA.svg", width = 9, height = 9)
ggsave(plot = bar.ind.NPA, "Enrich_plots/pdf/bar_ind_NPA.pdf", width = 9, height = 9)
ggsave(plot = bar.ind.NPA, "Enrich_plots/png/bar_ind_NPA.png", width = 9, height = 9)


ind.all.NPA.HPA <- ind.all.NPA.HPA[order(ind.all.NPA.HPA$precision),]
ind.all.NPA.HPA$term_name <- str_to_title(ind.all.NPA.HPA$term_name)
ind.all.NPA.HPA$term_name <- factor(ind.all.NPA.HPA$term_name, 
                                levels = ind.all.NPA.HPA$term_name)

bar.NPA.HPA <- ggplot(ind.all.NPA.HPA, aes(precision,term_name)) +
  geom_point(aes(fill= p_value, size= intersection_size), shape=21) +
  scale_fill_gradient(low="wheat", high= "olivedrab4",
                              name="FDR") +
  scale_size_continuous(breaks = c(5,10,15),
                        limits = c(0,15),
                        name = "Observed Gene Count") +
  xlim(0.05,0.12) +
  xlab("Gene Ratio") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 7))

bar.NPA.HPA
ggsave(plot = bar.NPA.HPA, "Enrich_plots/svg/bar_NPA_HPA.svg", width = 9, height = 9)
ggsave(plot = bar.NPA.HPA, "Enrich_plots/pdf/bar_NPA_HPA.pdf", width = 9, height = 9)
ggsave(plot = bar.NPA.HPA, "Enrich_plots/png/bar_NPA_HPA.png", width = 9, height = 9)


```


### TA 

```{r}

ind.all.TA <- r.tabl.all.TA[!r.tabl.all.TA$term_id %in% shared.ids,]
ind.all.TA <- ind.all.TA[,c(3,6:7,10:11)]
# Remove Transcription factors (we'll analyze them later)
ind.all.TA <- ind.all.TA[!ind.all.TA$source=="TF",]
ind.all.TA.HPA <- ind.all.TA[ind.all.TA$source=="HPA",]
# Trim to get only the top 40 hits
ind.all.TA <- ind.all.TA[!ind.all.TA$source=="HPA",]
ind.all.TA <- ind.all.TA[order(desc(ind.all.TA$p_value)),]
ind.all.TA <- ind.all.TA[253:292,]
ind.all.TA <- ind.all.TA[order(ind.all.TA$precision),]
ind.all.TA$source <- factor(ind.all.TA$source, levels = c("GO:BP","GO:CC",
                                                            "GO:MF","REAC",
                                                            "KEGG"))

ind.all.TA$term_name <- str_to_title(ind.all.TA$term_name)
ind.all.TA$term_name[17] <- "Lysosome (II)"
ind.all.TA$term_name <- factor(ind.all.TA$term_name, 
                                levels = ind.all.TA$term_name)

bar.ind.TA <- ggplot(ind.all.TA, aes(precision,term_name)) +
  geom_point(aes(fill= p_value, size= intersection_size,
                 color=source), shape=21, stroke=1.2) + 
  scale_color_manual(labels = c("GO: Biogical Process", 
                               "GO: Cellular Component",
                               "GO: Mollecular Function",
                               "REACTOME Pathway",
                               "KEGG database"
                               ),
                    values=c("#648FFF", "#785EF0",
                             "#DC267F","#A50D3E",
                             "#484661"),
                    name="Source Database") +
  scale_fill_gradient(low="wheat", high= "olivedrab4",
                              name="FDR") +
  scale_size_continuous(breaks = c(50,100,200,300),
                        limits = c(0,450),
                        name = "Observed Gene Count") +
  xlim(0,0.90) +
  xlab("Gene Ratio") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 7))

bar.ind.TA
ggsave(plot = bar.ind.TA, "Enrich_plots/svg/bar_ind_TA.svg", width = 9, height = 9)
ggsave(plot = bar.ind.TA, "Enrich_plots/pdf/bar_ind_TA.pdf", width = 9, height = 9)
ggsave(plot = bar.ind.TA, "Enrich_plots/png/bar_ind_TA.png", width = 9, height = 9)


ind.all.TA.HPA.lung <- ind.all.TA.HPA[ind.all.TA.HPA$term_name %in% 
                                ind.all.TA.HPA$term_name[startsWith(ind.all.TA.HPA$term_name,
                                                                    "lung")],]
ind.all.TA.HPA.bronch <- ind.all.TA.HPA[ind.all.TA.HPA$term_name=="bronchus",]
ind.TA.lungs <- rbind(ind.all.TA.HPA.bronch, ind.all.TA.HPA.lung)
ind.TA.lungs <- ind.TA.lungs[order(ind.TA.lungs$precision),]
ind.TA.lungs$term_name <- str_to_title(ind.TA.lungs$term_name)
ind.TA.lungs$term_name <- factor(ind.TA.lungs$term_name, 
                                levels = ind.TA.lungs$term_name)

bar.TA.lungs <- ggplot(ind.TA.lungs, aes(precision,term_name)) +
  geom_point(aes(fill= p_value, size= intersection_size), shape=21) +
  scale_fill_gradient(low="wheat", high= "olivedrab4",
                              name="FDR") +
  scale_size_continuous(breaks = c(50,60,80,90,100),
                        limits = c(0,322),
                        name = "Observed Gene Count") +
  xlim(0,0.85) +
  xlab("Gene Ratio") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 7))

bar.TA.lungs
ggsave(plot = bar.TA.lungs, "Enrich_plots/svg/bar_TA_lungs.svg", width = 9, height = 9)
ggsave(plot = bar.TA.lungs, "Enrich_plots/pdf/bar_TA_lungs.pdf", width = 9, height = 9)
ggsave(plot = bar.TA.lungs, "Enrich_plots/png/bar_TA_lungs.png", width = 9, height = 9)

```



# Enrichment for shared proteins

Created common proteins data set

```{r}

enrich.common <- rownames(noNA.prot)

```

Perform enrichment

```{r}
comm.enr <- gost(query = enrich.common, 
                organism = "hsapiens", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "fdr", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = NULL, as_short_link = FALSE)

gostplot(comm.enr, capped = TRUE, interactive = TRUE)

svglite::svglite("Figures/enrich_all_nogroup.svg")
gostplot(comm.enr, capped = TRUE, interactive = F)
dev.off()

res.comm.enr <- comm.enr$result
length(rownames(res.comm.enr[res.comm.enr$source=="GO:BP",]))
length(rownames(res.comm.enr[res.comm.enr$source=="GO:CC",]))
length(rownames(res.comm.enr[res.comm.enr$source=="GO:MF",]))
length(rownames(res.comm.enr[res.comm.enr$source=="KEGG",]))
length(rownames(res.comm.enr[res.comm.enr$source=="REAC",]))
length(rownames(res.comm.enr[res.comm.enr$source=="HPA",]))
length(rownames(res.comm.enr[res.comm.enr$source=="HP",]))
length(rownames(res.comm.enr[res.comm.enr$source=="WP",]))
length(rownames(res.comm.enr[res.comm.enr$source=="CORUM",]))
length(rownames(res.comm.enr[res.comm.enr$source=="TF",]))
length(rownames(res.comm.enr[res.comm.enr$source=="MIRNA",]))


```

Enrichment: LUNG

```{r}
res.comm.enr.HPA <- res.comm.enr[res.comm.enr$source=="HPA",]
comm.enr.lung <- res.comm.enr.HPA[res.comm.enr.HPA$term_name %in%
                 res.comm.enr.HPA$term_name[startsWith(res.comm.enr.HPA$term_name,
                                                      "lung")],]
comm.enr.bronch <- res.comm.enr.HPA[res.comm.enr.HPA$term_name %in%
                 res.comm.enr.HPA$term_name[startsWith(res.comm.enr.HPA$term_name,
                                                      "bronchus")],]
comm.enr.lungs <- rbind(comm.enr.lung,comm.enr.bronch)
comm.enr.lungs <- comm.enr.lungs[,c(3,6:7,10:11)]
comm.enr.lungs$term_name <- str_to_title(comm.enr.lungs$term_name)
comm.enr.lungs <- comm.enr.lungs[order(comm.enr.lungs$precision),]
comm.enr.lungs$term_name <- factor(comm.enr.lungs$term_name, 
                                levels = comm.enr.lungs$term_name)
write_xlsx(x = comm.enr.lungs[,c(5,4,2,3,1)], "Tables/Table_1.xlsx")


bar.comm.lungs <- ggplot(comm.enr.lungs, aes(precision,term_name)) +
  geom_point(aes(fill= p_value, size= intersection_size), shape=21) +
  scale_fill_gradient(low="wheat", high= "olivedrab4",
                              name="FDR") +
  scale_size_continuous(breaks = c(30,80,100,200,300,400),
                        limits = c(0,400),
                        name = "Observed Gene Count") +
  xlim(0,0.80) +
  xlab("Gene Ratio") +
  ylab("Enrichment terms") +
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1,
                                   size = 7))

bar.comm.lungs
ggsave(plot = bar.comm.lungs, "Enrich_plots/svg/bar_comm_lungs.svg", width = 9, height = 9)
ggsave(plot = bar.comm.lungs, "Enrich_plots/pdf/bar_comm_lungs.pdf", width = 9, height = 9)
ggsave(plot = bar.comm.lungs, "Enrich_plots/png/bar_comm_lungs.png", width = 9, height = 9)


```

# Enrichment: all (except HPA and TFs) PAUSED at the moment

```{r}

res.comm.enr.clean <- res.comm.enr[!res.comm.enr$source=="TF" &
                                     !res.comm.enr$source=="HPA",]

```
